---
title: "Kelp depth project"
output: html_notebook
---
#installing packages
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager") #need
BiocManager::install("decontam")
install.packages("devtools") #Installs devtools (if not already installed)
devtools::install_github("donaldtmcknight/microDecon",force=TRUE) #Installs microDecon
source("https://raw.githubusercontent.com/joey711/phyloseq/master/inst/scripts/installer.R",
       local = TRUE)
install_phyloseq(branch = "devel")
install.packages("ggplot2")
install.packages("vegan")
install.packages("BiocManager")
BiocManager::install("DESeq2")
install.packages("plotly")
install.packages("optparse")
install.packages("ggrepel")
install.packages("cluster")
install.packages("biomformat")
install.packages("gridBase")
library(devtools)
devtools::install_github("gauravsk/ranacapa")
install.packages("tidyr")
install.packages("emmeans")
install.packages("janitor")
install.packages("broom")
install.packages("treemapify")
install.packages("RColorBrewer")
```
#loading libraries
```{r}
library(dplyr)
library(tibble)
library(reshape2) #not used
library(phyloseq)
library(ggplot2)
library(vegan)
library(plotly) #not used
library(optparse) #not used
library(ggrepel)
library(cluster)
library(biomformat)
library('gridBase') #didn't use i think
library(ranacapa)
library(decontam)
library(tidyr)
library(emmeans)
library(janitor)
library(treemapify)
library(RColorBrewer)
```
#creating 12S post-decontam paths
```{r}
#setting working directory
setwd("~/Maura's Kelp Project/Kelp R stuff")
#12S metadata
metadata_12S <- read.csv("C:/Users/longc/OneDrive/Documents/Maura's Kelp Project/Kelp R stuff/rare_12S_meta.csv")
metadata_12S %>% 
  remove_rownames %>% 
  column_to_rownames(var="Sample_Acronym")-> metadata_12S
metadata_12S <- metadata_12S %>% mutate_all(na_if,"")
metadata_12S <- sample_data(metadata_12S)

#12S OTU
otu_12S <- read.csv("C:/Users/longc/OneDrive/Documents/Maura's Kelp Project/Kelp R stuff/rare_12S_otu.csv")

#converts first column into row names
otu_12S %>% 
  remove_rownames %>% 
  column_to_rownames(var="ESVId")-> otu_12S

otu_12S <- as.matrix(otu_12S, ncol=28)

class(otu_12S) <- "numeric"

otu_12S <- otu_table(otu_12S,taxa_are_rows=TRUE) #coercing it into a otu table

#12S tax
tax_12S <- read.csv("C:/Users/longc/OneDrive/Documents/Maura's Kelp Project/Kelp R stuff/rare_12S_tax.csv")
tax_12S %>% 
  remove_rownames %>% 
  column_to_rownames(var="ESVId")-> tax_12S

tax_12S <- as.matrix(tax_12S)
tax_12S <- tax_table(tax_12S)

#creating 12S phyloseq object
physeq_obj <- phyloseq(otu_12S,tax_12S,metadata_12S)
physeq_obj
```
#creating COI post-decontam paths
```{r}
metadata_COI <- read.csv("C:/Users/longc/OneDrive/Documents/Maura's Kelp Project/Kelp R stuff/rare_COI_meta.csv")
#COI metadata
metadata_COI %>% 
  remove_rownames %>% 
  column_to_rownames(var="Sample_Acronym")-> metadata_COI
metadata_COI <- sample_data(metadata_COI)
view(metadata_COI)
#COI OTU
otu_COI <- read.csv("C:/Users/longc/OneDrive/Documents/Maura's Kelp Project/Kelp R stuff/rare_COI_OTU.csv")
#converts first column into row names
otu_COI %>% 
  remove_rownames %>% 
  column_to_rownames(var="ESVId")-> otu_COI

otu_COI <- as.matrix(otu_COI, ncol=28)

class(otu_COI) <- "numeric"

otu_COI <- otu_table(otu_COI,taxa_are_rows=TRUE) #coercing it into a otu table

#COI tax
tax_COI <- read.csv("C:/Users/longc/OneDrive/Documents/Maura's Kelp Project/Kelp R stuff/rare_COI_tax.csv")
tax_COI %>% 
  remove_rownames %>% 
  column_to_rownames(var="ESVId")-> tax_COI
tax_COI <- as.matrix(tax_COI)
tax_COI <- tax_table(tax_COI)

#creating COI phyloseq object
coi_physeq_obj <- phyloseq(otu_COI,tax_COI,metadata_COI)
coi_physeq_obj

```
#phylo to vegan
```{r}
#converting 12S phyloseq object into something that can be analyzed by vegan functions
vegan_12S <- vegan_otu(physeq_obj)

#creating dataframe from metadata
sampledf <-   data.frame(sample_data(physeq_obj))
#creating Bray-curtis dissimilarity matrix
d_12S <- vegdist(vegan_12S, method="bray") 

#converting CO1 phyloseq object into something that can be analyzed by vegan functions
vegan_coi <- vegan_otu(coi_physeq_obj)

#creating Bray-Curtis dissimilarity matrix
d_COI <- vegdist(vegan_coi, method="bray") 

#creating dataframe from COI metadata table
sampledf2 <- data.frame(sample_data(coi_physeq_obj))
```
#NMDS PLOTS 12S
```{r}
physeq_ord <- ordinate(physeq_obj,"NMDS","bray")
#creating NMDS plot by depth
p.depth<- plot_ordination(physeq_obj,physeq_ord,color="Depth_Class")+stat_ellipse(geom="path")+ggtitle("No significant variation by depth for fish (12S) community assemblages ")+theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(plot.title = element_text(size=10))
p.depth
ggsave("12S_NMDS_depth.png")

#creating NMDS plot by region
p.region <- plot_ordination(physeq_obj,physeq_ord,color="Region")+ggtitle("Significant variation by region for fish (12S) community assemblages ")+stat_ellipse()+theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(plot.title = element_text(size=15))
p.region
ggsave("12S_NMDS_Region.png")
```
#NMDS PLOTS COI
```{r}
physeq_ord <- ordinate(coi_physeq_obj,"NMDS","bray")

#creating NMDS plot by depth
p.depth<- plot_ordination(coi_physeq_obj,physeq_ord,color="Depth_Class")+stat_ellipse(geom="path")+ggtitle("No significant variation by depth for invertebrates & algae (COI) community assemblages ")+theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(plot.title = element_text(size=10))
p.depth
ggsave("COI_NMDS_Depth.png")

#creating NMDS plot by region
p.region <- plot_ordination(coi_physeq_obj,physeq_ord,color="Region")+ggtitle("Significant variation by region for invertebrates & algae (COI) community assemblages ")+stat_ellipse()+theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(plot.title = element_text(size=8))
p.region
ggsave("COI_NMDS_Region.png")
```
#12SPERMANOVA
```{R}
#Model 1
broom::tidy(adonis(vegan_12S~sampledf$Depth_Class+sampledf$Region+sampledf$Depth_Class*sampledf$Region)$aov.tab)#depth around 0.1, region 0.003, interaction not sig

#Model 1 pairwise comparisons by depth
pairwise_adonis(vegan_12S,sampledf$Depth_Class,sim_method="bray",p_adjust_m = "fdr") #nothing

#Model 1 pairwise comparisons by region
pairwise_adonis(vegan_12S,sampledf$Region,sim_method="bray",p_adjust_m="fdr") #catalina is diff from the other 2

#Homogeneity of dispersions test
broom::tidy(TukeyHSD(betadisper(d_12S,getElement(sampledf,"Region")))) #not sig

#Model 2
broom::tidy(adonis(vegan_12S~sampledf$dissolved_oxygen+sampledf$nitrate+sampledf$chlorophyll+sampledf$Temp_9_m+sampledf$primary_productivity+sampledf$calcite+sampledf$light)$aov.tab) #chlorophyll = p=0.029

#Final model
broom::tidy(adonis(vegan_12S~sampledf$Region+sampledf$chlorophyll+sampledf$Region*sampledf$chlorophyll)$aov.tab) #region masks the effect of chlorophyll
```
##tree map
```{r}
#Model 1 tree map
m1 <- adonis(vegan_12S~sampledf$Depth_Class+sampledf$Region+sampledf$Depth_Class*sampledf$Region)$aov.tab#depth around 0.1, region 0.003, interaction not sig
q<-as.data.frame(m1)
row.names(q)<-c("Depth","Region","Sample","Residuals","Total") # must match names correctly
q<-data.frame(row.names(q), q)
names(q)[1]<-"Level"
q<-q[-which(row.names(q)=="Total"),]
q$Level<-paste(q$Level, "\n",
               round(q$R2,2), " (", q$Pr..F., ")", 
               sep="")
#Plot Figure 

q0.plot <- ggplot(q, aes(area = R2, label=Level))+
  geom_treemap(fill=brewer.pal(4, "Blues")[1:4], color=1, alpha=.8)+
  geom_treemap_text(colour = "black", place = "centre", grow = FALSE, size=10)+
  ggtitle("12S Apportioned Variance")+
  theme(plot.background = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))

q0.plot
ggsave("12S_m1.png")
```

```{r}
#Model 2 tree map
m2 <- adonis(vegan_12S~sampledf$dissolved_oxygen+sampledf$nitrate+sampledf$chlorophyll+sampledf$Temp_9_m+sampledf$primary_productivity+sampledf$calcite+sampledf$light)$aov.tab
q<-as.data.frame(m2)
row.names(q)<-c("Dissolved oxygen","nitrate","chlorophyll-a","temperature", "Primary Productivity", "calcite", "light", "Residuals", "Total") # must match names correctly
q<-data.frame(row.names(q), q)
names(q)[1]<-"Level"
q<-q[-which(row.names(q)=="Total"),]
q$Level<-paste(q$Level, "\n",
               round(q$R2,2), " (", q$Pr..F., ")", 
               sep="")
#Plot Figure 
qplot <- ggplot(q, aes(area = R2, label=Level))+
  geom_treemap(fill=brewer.pal(8, "Blues")[1:8], color=1, alpha=.8)+
  geom_treemap_text(colour = "black", place = "centre", grow = FALSE, size=13)+
  ggtitle("12S Apportioned Variance")+
  theme(plot.background = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
qplot
ggsave("12S_m2.png")
```

```{r}
#Final model tree map
m3 <- adonis(vegan_12S~sampledf$Region+sampledf$chlorophyll+sampledf$Region*sampledf$chlorophyll)$aov.tab
q<-as.data.frame(m3)
row.names(q)<-c("Region","chlorophyll-a","chlorophyll-a by region", "Residuals", "Total") # must match names correctly
q<-data.frame(row.names(q), q)
names(q)[1]<-"Level"
q<-q[-which(row.names(q)=="Total"),]
q$Level<-paste(q$Level, "\n",
               round(q$R2,2), " (", q$Pr..F., ")", 
               sep="")
#Plot Figure 
qplot <- ggplot(q, aes(area = R2, label=Level))+
  geom_treemap(fill=brewer.pal(4, "Blues")[1:4], color=1, alpha=.8)+
  geom_treemap_text(colour = "black", place = "centre", grow = FALSE, size=10)+
  ggtitle("12S Apportioned Variance")+
  theme(plot.background = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
ggsave("12S_m3.png")
```

#COI PERMANOVA
```{r}
#Model 1 PERMANOVA
broom::tidy(adonis(vegan_coi~sampledf2$Depth_Class+sampledf2$Region+sampledf2$Depth_Class*sampledf2$Region, permutations=10000)$aov.tab) #depth not sig, region sig, interaction not sig

#pairwise PERMANOVA by depth
pairwise_adonis(vegan_coi,sampledf2$Depth_Class,sim_method="bray",p_adjust_m="fdr") #not sig

#pairwise PERMANOVA by region
pairwise_adonis(vegan_coi,sampledf2$Region,sim_method="bray",p_adjust_m="fdr") 

#Model 2 PERMANOVA
broom::tidy(adonis(vegan_coi~sampledf2$dissolved_oxygen+sampledf2$nitrate+sampledf2$chlorophyll+sampledf2$Temp_9_m+sampledf2$light+sampledf2$primary_productivity+sampledf2$calcite)$aov.tab)

#Final model PERMANOVA
broom::tidy(adonis(vegan_coi~sampledf2$Region+sampledf2$dissolved_oxygen+sampledf2$chlorophyll+sampledf2$Region*sampledf2$chlorophyll+sampledf2$Region*sampledf2$dissolved_oxygen)$aov.tab)

#Homogeneity of dispersions test
broom::tidy(TukeyHSD(betadisper(d_COI,getElement(sampledf2,"Region"))))
```
##tree map
```{r}
#Model 1 tree map
m1 <- adonis(vegan_coi~sampledf2$Depth_Class+sampledf2$Region+sampledf2$Depth_Class*sampledf2$Region)$aov.tab#depth around 0.1, region 0.003, interaction not sig
q<-as.data.frame(m1)
row.names(q)<-c("Depth","Region","Sample","Residuals","Total") # must match names correctly
q<-data.frame(row.names(q), q)
names(q)[1]<-"Level"
q<-q[-which(row.names(q)=="Total"),]
q$Level<-paste(q$Level, "\n",
               round(q$R2,2), " (", q$Pr..F., ")", 
               sep="")

qplot <- ggplot(q, aes(area = R2, label=Level))+
  geom_treemap(fill=brewer.pal(4, "Blues")[1:4], color=1, alpha=.8)+
  geom_treemap_text(colour = "black", place = "centre", grow = FALSE, size=10)+
  ggtitle("COI Apportioned Variance")+
  theme(plot.background = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
qplot
ggsave("coi_m1.png")
```
```{r}
#Model 2 tree map
m2 <- adonis(vegan_coi~sampledf2$dissolved_oxygen+sampledf2$nitrate+sampledf2$chlorophyll+sampledf2$Temp_9_m+sampledf2$primary_productivity+sampledf2$calcite+sampledf2$light)$aov.tab
broom::tidy(m2)
q<-as.data.frame(m2)
row.names(q)<-c("Dissolved oxygen","nitrate","chlorophyll-a","temperature", "Primary Productivity", "calcite", "light", "Residuals", "Total") # must match names correctly
q<-data.frame(row.names(q), q)
names(q)[1]<-"Level"
q<-q[-which(row.names(q)=="Total"),]
q$Level<-paste(q$Level, "\n",
               round(q$R2,2), " (", q$Pr..F., ")", 
               sep="")
#Plot Figure 
qplot <- ggplot(q, aes(area = R2, label=Level))+
  geom_treemap(fill=brewer.pal(8, "Blues")[1:8], color=1, alpha=.8)+
  geom_treemap_text(colour = "black", place = "centre", grow = FALSE, size=7)+
  ggtitle("CO1 Apportioned Variance")+
  theme(plot.background = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
qplot
ggsave("coi_m2.png")
```
```{R}
#Final model tree map
m3 <- adonis(vegan_coi~sampledf2$Region+sampledf2$chlorophyll+sampledf2$dissolved_oxygen+sampledf2$Region*sampledf2$chlorophyll+sampledf2$Region*sampledf2$dissolved_oxygen)$aov.tab
broom::tidy(m3)
q<-as.data.frame(m3)
row.names(q)<-c("Region","chlorophyll-a", "dissolved oxygen", "chlorophyll-a by region", "dissolved oxygen by region", "Residuals", "Total") # must match names correctly
q<-data.frame(row.names(q), q)
names(q)[1]<-"Level"
q<-q[-which(row.names(q)=="Total"),]
q$Level<-paste(q$Level, "\n",
               round(q$R2,2), " (", q$Pr..F., ")", 
               sep="")
#Plot Figure 
ggplot(q, aes(area = R2, label=Level))+
  geom_treemap(fill=brewer.pal(6, "Blues")[1:6], color=1, alpha=.8)+
  geom_treemap_text(colour = "black", place = "centre", grow = FALSE, size=10)+
  ggtitle("COI Apportioned Variance")+
  theme(plot.background = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
ggsave("coi_m3.png")
```
#CCA
```{r}
#The full CCA visualization was buggy and didn't let the names of the taxa show, only the ESV IDs, but we cross referenced with the taxonomy table
```
##12S Region
```{r}
cap_ord <- ordinate(
  physeq = physeq_obj, 
  method = "CAP",
  distance = d_12S,
  formula = ~Region)
#Results of model
anova(cap_ord)
```

```{r}
# CAP plot of 12S predictor taxa
sample_data(physeq_obj)$Region = factor(sample_data(physeq_obj)$Region, levels = c("Catalina","Malibu","Palos_Verdes"))
get_variable(physeq_obj,"Region")
cap_plot <- plot_ordination(
  physeq = physeq_obj, 
  ordination = cap_ord, 
  color = "Region", 
  axes = c(1,2) #what this mean
) + 
  aes(shape = Region) + 
  geom_point(aes(colour = Region), size = 4) + 
  geom_point(colour = "grey90", size = 1.5)
cap_plot
```

```{r}
# Now add the environmental variables as arrows
arrowmat <- vegan::scores(cap_ord, display = "bp")
# Add labels, make a data.frame #WHAT
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)
# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = 0.92 * CAP1, 
                 y = 1.3 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))
#Hulls for each polygon

test <- vegan::scores(cap_ord, display = "sites")
data.scores_1 <- as.data.frame(scores(test))  

#Using the scores function from vegan to extract the site scores and 
# convert to a data.frame
data.scores_1$sites <- sampledf$Region
# create a column of site names, from the rownames of data.scores

#Add location data 
grp <- sampledf$Region
data.scores_1$grp <- grp  #  add the grp variable created earlier
```

```{r}
##### Calculate Shape Around Points
grp.pv <- data.scores_1[data.scores_1$sites == "Palos_Verdes", ][chull(data.scores_1[data.scores_1$sites == "Palos_Verdes", c("CAP1", "CAP2")]),] #it's only extracting the first three sites
grp.malibu <- data.scores_1[data.scores_1$grp == "Malibu", ][chull(data.scores_1[data.scores_1$grp == "Malibu", c("CAP1", "CAP2")]), ] #it's only extracting the first three sites
grp.catalina <- data.scores_1[data.scores_1$grp == "Catalina", ][chull(data.scores_1[data.scores_1$grp == "Catalina", c("CAP1", "CAP2")]), ] #it's only extracting the first three sites
hull.data <- rbind(grp.pv,grp.malibu,grp.catalina)
colnames(hull.data)[3] <- c("Region")
```

```{r}


# Make a new graphic for the constrained ordination

cap_1 <- cap_plot +  
  theme_bw(base_size = 18)+
geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "grey", 
    arrow = arrowhead) +
  geom_text(
    mapping = label_map, 
    size = 6,  
    data = arrowdf, 
    show.legend = FALSE
  ) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2))+
  geom_polygon(data=hull.data,aes(x=CAP1,y=CAP2,fill=grp, group=grp),alpha=0.30) +
    ggtitle("Constrained Analysis of Principle Components")

cap_1

```
###predictor species

```{r}
#Retain species scores
cap_12S <- capscale(vegan_12S ~ Region, data=sampledf, dist="bray") 
sppscores(cap_12S) <- vegan_12S #adding species scores from vegan_12S to cap_12S
```

```{r}
as.data.frame(vegan::scores(cap_12S, display="species")) %>%
  rownames_to_column(var = "sample") %>% 
  as_tibble %>%
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) -> cap_12S_species_distances 

cap_12S_species_distances %>% arrange(dist) %>% ggplot(., aes(y=dist, x=reorder(sample,-dist)))+geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
ggplot(data=cap_12S_species_distances, aes(cap_12S_species_distances$dist)) + geom_histogram()
```

```{r}
cap_12S_species_distances %>% 
  top_n(7, dist) -> top_species
top_species <- as.data.frame(top_species)
rownames(top_species) <- top_species$sample

# Now add the environmental variables as arrows
arrowmat <- top_species #up to CAP2

#Rename row names to clean up plot
#it's not calling the names!!! ahhhh
view(arrowmat)
rownames(top_species) %>% as.data.frame() -> namers
colnames(namers) <- c("path")
namers %>% separate(path, c("Domain","Phylum","Class","Order","Family","Genus","Species"), sep=";") %>% #
  mutate(., name = ifelse(Species == "", Genus, Species)) %>% 
  mutate(., name = ifelse(name == "", Family, name)) %>% 
  mutate(., name = ifelse(name == "", Order, name)) %>% 
  mutate(., name = ifelse(name == "", Class, name)) %>% 
  mutate(., name = ifelse(name == "", Phylum, name))->namers
```
##don't run until you figure out why it's not assigning names
```{R}
rownames(arrowmat) <- NULL
rownames(arrowmat) <- namers$name #still up to CAP2
```

```{r}

# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat) #up to CAP2

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = 0.75 * CAP1, 
                 y = 0.75 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

#Hulls for each polygon


test.2 <- as.data.frame(cap_12S$CCA$wa) 

data.scores_1 <- as.data.frame(scores(test.2)) 

data.scores_1$Region <- sampledf$Region
# create a column of region names, from the rownames of data.scores
head(data.scores_1) 


grp <- sampledf$Region
data.scores_1$grp <-grp  #  add the grp variable created earlier
  
#### Calculate Convex Hull Shape Around Points 


grp.pv <- data.scores_1[data.scores_1$sites == "Palos_Verdes", ][chull(data.scores_1[data.scores_1$sites == "Palos_Verdes", c("CAP1", "CAP2")]),] #it's only extracting the first three sites
grp.malibu <- data.scores_1[data.scores_1$grp == "Malibu", ][chull(data.scores_1[data.scores_1$grp == "Malibu", c("CAP1", "CAP2")]), ] #it's only extracting the first three sites
grp.catalina <- data.scores_1[data.scores_1$grp == "Catalina", ][chull(data.scores_1[data.scores_1$grp == "Catalina", c("CAP1", "CAP2")]), ] #it's only extracting the first three sites
hull.data <- rbind(grp.pv,grp.malibu,grp.catalina)
 
# Plot CAP 
data_cap <- as.data.frame(cap_12S$CCA$wa)%>% #forcing this into a dataframe?
  rownames_to_column(var = "sample") %>% 
  as_tibble() %>% 
  separate(sample, into = c("Region")) 
#had to add data to figure it out

ggplot(data=data_cap,aes(x=CAP1, y=CAP2)) + geom_point() + 
  geom_segment(mapping = arrow_map, size = .5, data = arrowdf,color = "grey",arrow = arrowhead)+
  geom_text(mapping = label_map, size = 5, data = arrowdf, show.legend = FALSE, position=position_jitter(width=0.05,height=0.08))+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2))+
  geom_polygon(data=hull.data,aes(x=CAP1,y=CAP2,fill=grp, group=grp),alpha=0.30) + 
   ggtitle("Predictor 12S taxa driving regional variation",)+
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        axis.text=element_text(size=20),
        axis.title=element_text(size=24,face="bold"),
        legend.title=element_text(size=20),
        legend.text=element_text(size=18),
        plot.subtitle = element_text(size=20),
        legend.spacing.y = unit(0.5, 'cm'), 
        legend.key = element_rect(size = 10),
        legend.key.size = unit(2, 'lines'))
 
```
##COI Region
```{r}
cap_ord <- ordinate(
  physeq = coi_physeq_obj, 
  method = "CAP",
  distance = d_COI,
  formula = ~Region)
#Results of model
anova(cap_ord)
```

```{r}
# CAP plot of environmental predictor variables
sample_data(coi_physeq_obj)$Region = factor(sample_data(coi_physeq_obj)$Region, levels = c("Catalina","Malibu","Palos_Verdes"))
get_variable(physeq_obj,"Region")
cap_plot <- plot_ordination(
  physeq = coi_physeq_obj, 
  ordination = cap_ord, 
  color = "Region", 
  axes = c(1,2) #what this mean
) + 
  aes(shape = Region) + 
  geom_point(aes(colour = Region), size = 4) + 
  geom_point(colour = "grey90", size = 1.5)
cap_plot
```

```{r}
# Now add the environmental variables as arrows
arrowmat <- vegan::scores(cap_ord, display = "bp")
# Add labels, make a data.frame #WHAT
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)
# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = 0.92 * CAP1, 
                 y = 1.3 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))
#Hulls for each polygon

test <- vegan::scores(cap_ord, display = "sites")
data.scores_1 <- as.data.frame(scores(test))  

#Using the scores function from vegan to extract the site scores and 
# convert to a data.frame
data.scores_1$sites <- sampledf2$Region
# create a column of site names, from the rownames of data.scores

#Add location data 
grp <- sampledf2$Region
data.scores_1$grp <- grp  #  add the grp variable created earlier
```

```{r}
##### Calculate Shape Around Points
grp.pv <- data.scores_1[data.scores_1$sites == "Palos_Verdes", ][chull(data.scores_1[data.scores_1$sites == "Palos_Verdes", c("CAP1", "CAP2")]),] #it's only extracting the first three sites
grp.malibu <- data.scores_1[data.scores_1$grp == "Malibu", ][chull(data.scores_1[data.scores_1$grp == "Malibu", c("CAP1", "CAP2")]), ] #it's only extracting the first three sites
grp.catalina <- data.scores_1[data.scores_1$grp == "Catalina", ][chull(data.scores_1[data.scores_1$grp == "Catalina", c("CAP1", "CAP2")]), ] #it's only extracting the first three sites
hull.data <- rbind(grp.pv,grp.malibu,grp.catalina)
colnames(hull.data)[3] <- c("Region")
```

```{r}


# Make a new graphic for the constrained ordination

cap_1 <- cap_plot +  
  theme_bw(base_size = 18)+
geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "grey", 
    arrow = arrowhead) +
  geom_text(
    mapping = label_map, 
    size = 6,  
    data = arrowdf, 
    show.legend = FALSE
  ) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2))+
  geom_polygon(data=hull.data,aes(x=CAP1,y=CAP2,fill=grp, group=grp),alpha=0.30) +
    ggtitle("Constrained Analysis of Principle Components")

cap_1

```
###predictor species

```{r}
#Retain species scores
cap_COI <- capscale(vegan_coi ~ Region, data=sampledf2, dist="bray") 
sppscores(cap_COI) <- vegan_coi #adding species scores from vegan_12S to cap_12S
```

```{r}
as.data.frame(vegan::scores(cap_COI, display="species")) %>%
  rownames_to_column(var = "sample") %>% 
  as_tibble %>%
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) -> cap_COI_species_distances 

cap_COI_species_distances %>% arrange(dist) %>% ggplot(., aes(y=dist, x=reorder(sample,-dist)))+geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
ggplot(data=cap_COI_species_distances, aes(cap_COI_species_distances$dist)) + geom_histogram()
```

```{r}
cap_COI_species_distances %>% 
  top_n(7, dist) -> top_species
top_species <- as.data.frame(top_species)
rownames(top_species) <- top_species$sample

# Now add the environmental variables as arrows
arrowmat <- top_species #up to CAP2

#Rename row names to clean up plot
#it's not calling the names!!! ahhhh
view(arrowmat)
rownames(top_species) %>% as.data.frame() -> namers
colnames(namers) <- c("path")
namers %>% separate(path, c("Domain","Phylum","Class","Order","Family","Genus","Species"), sep=";") %>% #
  mutate(., name = ifelse(Species == "", Genus, Species)) %>% 
  mutate(., name = ifelse(name == "", Family, name)) %>% 
  mutate(., name = ifelse(name == "", Order, name)) %>% 
  mutate(., name = ifelse(name == "", Class, name)) %>% 
  mutate(., name = ifelse(name == "", Phylum, name))->namers
```
###don't run until you figure out why it's not assigning names
```{R}
rownames(arrowmat) <- NULL
rownames(arrowmat) <- namers$name #still up to CAP2
```

```{r}

# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat) #up to CAP2

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = 0.75 * CAP1, 
                 y = 0.75 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

#Hulls for each polygon


test.2 <- as.data.frame(cap_COI$CCA$wa) 

data.scores_1 <- as.data.frame(scores(test.2)) 

data.scores_1$Region <- sampledf2$Region
# create a column of region names, from the rownames of data.scores
head(data.scores_1) 


grp <- sampledf2$Region
data.scores_1$grp <-grp  #  add the grp variable created earlier
  
#### Calculate Convex Hull Shape Around Points 


grp.pv <- data.scores_1[data.scores_1$sites == "Palos_Verdes", ][chull(data.scores_1[data.scores_1$sites == "Palos_Verdes", c("CAP1", "CAP2")]),] #it's only extracting the first three sites
grp.malibu <- data.scores_1[data.scores_1$grp == "Malibu", ][chull(data.scores_1[data.scores_1$grp == "Malibu", c("CAP1", "CAP2")]), ] #it's only extracting the first three sites
grp.catalina <- data.scores_1[data.scores_1$grp == "Catalina", ][chull(data.scores_1[data.scores_1$grp == "Catalina", c("CAP1", "CAP2")]), ] #it's only extracting the first three sites
hull.data <- rbind(grp.pv,grp.malibu,grp.catalina)
 
# Plot CAP 
data_cap <- as.data.frame(cap_COI$CCA$wa)%>% #forcing this into a dataframe?
  rownames_to_column(var = "sample") %>% 
  as_tibble() %>% 
  separate(sample, into = c("Region")) 
#had to add data to figure it out

ggplot(data=data_cap,aes(x=CAP1, y=CAP2)) + geom_point() + 
  geom_segment(mapping = arrow_map, size = .5, data = arrowdf,color = "grey",arrow = arrowhead)+
  geom_text(mapping = label_map, size = 5, data = arrowdf, show.legend = FALSE, position=position_jitter(width=0.05,height=0.08))+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2))+
  geom_polygon(data=hull.data,aes(x=CAP1,y=CAP2,fill=grp, group=grp),alpha=0.30) + 
   ggtitle("Predictor COI taxa driving regional variation",)+
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        axis.text=element_text(size=20),
        axis.title=element_text(size=24,face="bold"),
        legend.title=element_text(size=20),
        legend.text=element_text(size=18),
        plot.subtitle = element_text(size=20),
        legend.spacing.y = unit(0.5, 'cm'), 
        legend.key = element_rect(size = 10),
        legend.key.size = unit(2, 'lines'))
 
```
#bar plots for visualization
##COI CCA taxa
```{r}
Micromonas_pusilla <- subset_taxa(coi_physeq_obj,Species=="Micromonas pusilla")
title="Distribution of Micromonas pusilla among samples"
p <- plot_bar(Micromonas_pusilla,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")

```
```{r}
Bathycoccus_prasinos <- subset_taxa(coi_physeq_obj,Species=="Bathycoccus prasinos")
title="Distribution of Bathycoccus prasinos among samples"
p <- plot_bar(Bathycoccus_prasinos,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```
```{r}
Sarsamphiascus_kawamurai  <- subset_taxa(coi_physeq_obj,Species=="Sarsamphiascus kawamurai")
title="Distribution of Sarsamphiascus kawamurai among samples"
p <- plot_bar(Sarsamphiascus_kawamurai ,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```
```{r}
haptoglossa  <- subset_taxa(coi_physeq_obj,Species=="Haptoglossa sp. DM2 4/12")
title="Distribution of Haptoglossa sp. DM2 4/12 among samples"
p <- plot_bar(haptoglossa ,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```
```{r}
Podocorynoides_minima   <- subset_taxa(coi_physeq_obj,Species=="Podocorynoides minima")
title="Distribution of Podocorynoides minima among samples"
p <- plot_bar(Podocorynoides_minima  ,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```
```{r}
Pseudevadne_tergestina <- subset_taxa(coi_physeq_obj,Species=="Pseudevadne tergestina")
title="Distribution of Pseudevadne tergestina among samples"
p <- plot_bar(Pseudevadne_tergestina  ,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```
```{r}
Evadne_nordmanni <- subset_taxa(coi_physeq_obj,Species=="Evadne nordmanni")
title="Distribution of Evadne nordmanni among samples"
p <- plot_bar(Evadne_nordmanni  ,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```
##12S CCA taxa
```{r}
senorita <- subset_taxa(physeq_obj,Species=="Oxyjulis californica")
title="Distribution of senorita among sites"
p <- plot_bar(senorita,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")

```

```{r}
kelp_bass <- subset_taxa(physeq_obj,Species=="Paralabrax clathratus")
title="Distribution of kelp bass among sites"
p <- plot_bar(kelp_bass,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```
```{r}
northern_anchovy <- subset_taxa(physeq_obj,Species=="Engraulis mordax")
title="Distribution of northern anchovy among sites"
p <- plot_bar(northern_anchovy,"Region","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```

```{r}
california_grunion <- subset_taxa(physeq_obj,Species=="Leuresthes tenuis")
title="Distribution of california grunion among sites"
p <- plot_bar(california_grunion,"Region","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```
```{r}
california_tonguefish <- subset_taxa(physeq_obj,Species=="Symphurus atricaudus")
title="Distribution of California tonguefish among sites"
p <- plot_bar(california_tonguefish,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```
```{r}
kelp_perch <- subset_taxa(physeq_obj,Species=="Brachyistius frenatus")
title="Distribution of kelp perch among samples"
p <- plot_bar(kelp_perch,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```
```{r}
california_sheephead <- subset_taxa(physeq_obj,Species=="Semicossyphus pulcher")
title="Distribution of California Sheephead among samples"
p <- plot_bar(california_sheephead,"Sample","Abundance","Species",title=title)
p+geom_bar(aes(fill=Region),stat="identity",position="stack")
```
#linear model (chlorophll ~ region)
```{r}
m1 <- lm(chlorophyll~Region,data=sampledf)
summary(m1)
anova(m1)#sig with temp average and temp no m
#maybe there is a temperature gradient, but just across all sites throughout the southern california bight, but not in between depths. 
```
